Authentication based on DRAM PUFs

### 2021-06-17 latexPaper/paper.tex
arrange references
update

@@

% Write abstract here
\begin{abstract}
One of the major security concerns regarding IoT is that of counterfeit or tampered hardware. To overcome this problem it is important to employ a hardware-intrinsic security mechanism for authentication of the device. While previous solutions like secure key storage attempt to solve this problem they only pass on the issue to the next stage. This paper outlines the use of Physically Unclonable Functions (PUFs) for hardware-intrinsic authentication. It summarizes well-known works on DRAM based PUFs and also presents known protocols and schemes to achieve authentication on embedded devices using PUFs.
\end{abstract}

@@

In practice most PUFs are characterized by challenges and responses. A challenge is the input given to the PUF. The PUF utilizes random manufacturing variations to convert the challenge to a response. The randomness makes the response unpredictabl. And because these variations cannot be controlled, even by the manufacturer, the response is unclonable on any other device. Thios way they attain their property of being \textbf{physically unclonable}.

@@

\textbf{Startup:} In this method, the startup state of the DRAM is used for unique fingerprinting. As mentioned by Tehranipoor et al. “DRAMs have seemingly random startup values”. This is because the bitlines on DRAM are precharged to vdd/2 which makes the sense amplifier equally likely to read a ‘1’ or ‘0’. But, due to the process variations of manufacturing most bits have either a bias towards a ‘1’ or a ‘0’. Hence, the startup values of the DRAM are randomized by the random process variations. It is noted though that some bits have a neutral skew/bias i.e. they randomly switch to ‘1’ or ‘0’ over multiple startups. These bits can be used for random number generation. This method is not suitable for authentication using the traditional CRP database approach, because startup values cannot be varied by writing the challenge to memory before startup.\cite{TEH15}

\textbf{Write Failures:} In this method, process variations affecting the write reliability are exploited. Under normal operation, the write enable signal activates the bit lines i.e. connects them to the data bus which then overrides the bit line voltage. As a result, the DRAM cells would be overwritten by the value on the data bus. For normal operation the duty cycle of the write signal is chosen such that all DRAM cells can successfully. But for PUF operation, the duty cycle is deliberately reduced which results in some DRAM cells being overwritten while others are not. If a cell has been overwritten or not depends on the manufacturing variations. The data written under PUF operation can be treated as the challenge and a subsequent read reveals the response affected by the variations. \cite{HASH}

@@

\section{Implementation Objectives}
In order to analyse the various presented methods measurements have to be made. Almost all surveyed papers use either hamming distance and fractional hamming distance or jaccard index for measuring the properties of the PUFs.\cite{jaccard} The usage of one over the other is debatable although Xiong et al claims jaccard index to be a better metric because it takes into consideration only the bits that flip as opposed to fractional hamming distance which weighs all bits of the CRP equally.

@@

\textit{Uniqueness.} Uniqueness in PUFs is measured by comparing responses for the same challenges on different PUF instances. Greater difference in responses indicates stronger uniqueness.  The comparison is measured using either hamming distance(needs to be high) or the jaccard index(needs to be low). All surveyed work proves the uniqueness through experimentation on PUF instances of different DRAM modules, Sutar et al. goes a step further to analyze the uniqueness of different PUF instances within the same DRAM module. Since the experiment conducted by Sutar et al. sparks interest the table below summarizes their results.

@@

\begin{figure}[!hbt]
    % Center the figure.
    \begin{center}
    % Include the eps file, scale it such that it's width equals the column width. You can also put width=8cm for example...
    \includegraphics[scale=0.5]{figs/enrollment.png}
    % Create a subtitle for the figure.
    \caption{Enrollment}
    % Define the label of the figure. It's good to use 'fig:title', so you know that the label belongs to a figure.
    \label{fig:enroll}
    \end{center}
\end{figure}

@@

\textit{One Way Authentication.} The flowchart in Fig.\ref{fig:one_way} depicts the one-way authentication process. In short, The server matches the device's remeasured response against the CRP database. If matches, the device is authenticated

\begin{figure}[!hbt]
    % Center the figure.
    \begin{center}
    % Include the eps file, scale it such that it's width equals the column width. You can also put width=8cm for example...
    \includegraphics[scale=0.5]{figs/authentication.png}
    % Create a subtitle for the figure.
    \caption{1-way Authentication}
    % Define the label of the figure. It's good to use 'fig:title', so you know that the label belongs to a figure.
    \label{fig:one_way}
    \end{center}
\end{figure}

\textit{Mutual Authentication.} This is an extension of the 1-way authentication which requires storage of CRPs in form of addresses of flipped bits rather than the bit pattern itself. The device mixes random addresses with the flipped addresses before sending the response to the server. This is done to test the knowledge of the server. The server verfies the device by counting the number of matched addresses and checking it against a threshold. The server then filters out the known flip addresses from the response and sends it to the device. The device verifies the server's response by matching it with its own response.

In both authentication protocols described above, it is important to note that:
\begin{itemize}
    \item A CRP is never used twice to prevent replay attacks. Hence a large CRP database or a strong PUF is required for continued authentication.
    \item For fast authentication it is critical to pick a CRP with the fastest possible parameters(low refresh pause interval) is picked. This requires knowledge of the PUF characteristics beforehand.
\end{itemize}

@@

\textit{Parameter Reconfigurability.} The main objective of parameter reconfigurability is to increase the size of the CRP database. Fot this different parameters that can change the response behaviour of the PUF have to be recognized

Sutar et al in their work show the reconfigurability of the refresh pause interval in order to obtain new challenge response pairs on the same PUF instance. A similar experiment is also conducted by Xiong et al and Schaller et al. In both these works, the lowest refresh pause interval was 20s. After the CRPs of a particular refresh pause interval have been exhausted the interval can be increased to higher one depending on the number of “new” bit flips. Both these works also propose increasing the size of the PUF instance, in order to maintain the low refresh pause interval at the cost of DRAM space exhaustion.

@@

With these parameters at disposal, one can reconfigure the PUF instance to generate new sets of CRP pairs and thereby increasing the size of the CRP database.

\textit{Characterization:} For proper usage of the reconfiguration parameters, the challenge response behaviour must be analysed before the authentication can take place. For achieving this the DRAM module has to be \textit{characterized}. In simple terms, characterization is the process of trying all kinds of combinations of parameters and observing the difference in responses.\cite{SUT18} After such an exhaustive experimentation the best suitable parameters are selected and the CRP database is generated accordingly.

@@

When a PUF cannot generate a large CRP database, it is used with a Fuzzy Extractor a.k.a Helper Data System \cite{HDS15} for secure key storage. A fuzzy extractor corrects errors in the PUF response \cite{SUT18} and enables regeneration of a cryptogrpahic key. This way the reproduced key can be used to setup a secure channel through symmetric protocols. Once a secure channel is established authentication can be done through asymmetric protocols. A general construction of a fuzzy extractor contains:

\begin{itemize}
    \item \textit{Enrollment Algorithm:} The enrollment algorithm takes as input a PUF response \(X\) and a random cryptographic key \(k\) and outputs helper data \(w\).
    \item \textit{Reconstruction Algorithm:}  The reconstruction algorithm takes as input a fresh and noisy measurement of the PUF response \(X'\) and helper data \(w\) and outputs a reconstructed key \(k'\). If the noise in the PUF response is within a threshold \( k' = k\)
\end{itemize}

@@

One of the earliest implementations of fuzzy extractors is the Code Offset Method. \cite{COM} COM uses a error correcting code at the core of its construction. The helper data \(w = X \oplus Enc(k)\), where \(Enc()\) is the encoding function of the error correcting code. The reconstructed key \(k' = Dec(X' \oplus w)\), where \(Dec()\) is the decoding function of the error correcting code. The important point to note here is that \(w\) plays the role of a syndrome. The error correcting code extracts the key \(k\) by correcting errors in \(X'\). This is clear when the formula is rewritten as \(k' = Dec(X' \oplus X \oplus Enc(k))\). Any remaining bits of the \(X' \oplus X\), if there are any, is corrected by \(Dec()\)

Using such a fuzzy extractor would also require sufficient entropy in the input \(X\). Schaller et al overcomes this problem by selecting equal number of \textbf{F} cells and \textbf{S} cells. The entropy in the input is required so that no information is leaked through the helper data \(w\).

Another version of the COM scheme, Syndrome-COM or SCOM,\cite{COM} leverages syndrome decoding to use the input \(X\) itself as the cryptographic key. In this scheme the helper data \(w = Syn(X)\), where \(Syn()\) is the syndrome function. The reconstructed key \(k' = SynDec(w \oplus Syn(X'))\). The expression \(w \oplus Syn(X')\) equals \(Syn(X \oplus X')\), i.e., the syndrome of the error vector and \(SynDec()\) is teh syndrom decoder which corrects the error.

@@

Other constructions, of HDSs make use of concatenation of two linear codes \(C_{1} = [n_{1}, k_{1}, d_{1}]\) \( c_{2} = [n_{2}, k_{2}, d_{2}]\).\cite{HDS15} In this construction, Code \(C_{1}\) needs to maintain a high entropy, which directly translates into high cardinality or high \(k_{1}\). And, Code \(C_{2}\) needs to have a high error correction rate, whcih directly transaltes to higher distance \(d_{2}\).

\section{Conclusion}
This paper highlights why and how PUFs fufill the demands of hardware-intrinsic authentication. It explains the well known methods of implementing PUFs on DRAMs and the advantages of DRAM PUFs. It outlines the objectives a PUF implementation has to meet in order to be fit for use in authentication. It summarizes a simple lightweight protocol that can meet strict resource constraints for embedded processors. It also dives into the details of using PUFs for secure key storage using helper data systems (HDS) and summarizes different constructions for HDS.

@@

    \bibitem{KELLER}
    C. Keller et al., "Dynamic memory-based physically unclonable function for the generation of unique identifiers and true random numbers," 2014 IEEE International Symposium on Circuits and Systems (ISCAS), 2014, pp. 2740-2743, doi: 10.1109/ISCAS.2014.6865740.

@@

    \bibitem{jaccard}
    Jaccard, P. (1901) étude Comparative de la distribuition florale dans une portion des Alpes et des Jura. Bulletin de la Société Vaudoise des Sciences Naturelles, 7, 547-579.

>>>>

\usepackage{algorithm,algorithmic}

@@

% Write abstract here
\begin{abstract}
One of the major security concerns regarding IoT is that of counterfeit or tampered hardware. To overcome this problem it is important to employ a hardware-intrinsic security mechanism for authentication of the device. While previous solutions like secure key storage attempt to solve this problem they only pass on the issue to the next stage. This paper outlines the use of Physically Unclonable Functions (PUFs) for hardware-intrinsic authentication. It summarizes well-known works on DRAM-based PUFs and also presents known protocols and schemes to achieve authentication on embedded devices using PUFs.
\end{abstract}

@@

In practice most PUFs are characterized by challenges and responses. A challenge is the input given to the PUF. The PUF utilizes random manufacturing variations to convert the challenge to a response. The randomness makes the response unpredictable. And because these variations cannot be controlled, even by the manufacturer, the response is unclonable on any other device. This way they attain their property of being \textbf{physically unclonable}.

@@

\textbf{Startup:} In this method, the startup state of the DRAM is used for unique fingerprinting. As mentioned by Tehranipoor et al. “DRAMs have seemingly random startup values”. This is because the bitlines on DRAM are precharged to vdd/2 which makes the sense amplifier equally likely to read a ‘1’ or ‘0’. But, due to the process variations of manufacturing most bits have either a bias towards a ‘1’ or a ‘0’. Hence, the startup values of the DRAM are randomized by the random process variations. It is noted though that some bits have a neutral skew/bias i.e. they randomly switch to ‘1’ or ‘0’ over multiple startups. These bits can be used for random number generation. \cite{TEH15}

\textbf{Write Failures:} In this method, process variations affecting the write reliability are exploited. Under normal operation, the write enable signal activates the bit lines i.e. connects them to the data bus which then overrides the bit line voltage. As a result, the DRAM cells would be overwritten by the value on the data bus. For normal operation the duty cycle of the write signal is chosen such that all DRAM cells can be successfully overwritten. But for PUF operation, the duty cycle is deliberately reduced which results in some DRAM cells being overwritten while others are not. If a cell has been overwritten or not depends on the manufacturing variations. The data written under PUF operation can be treated as the challenge and a subsequent read reveals the response affected by the variations. \cite{HASH}

@@

\section{Implementation Objectives}
In order to analyze the various presented methods, measurements have to be made. Almost all surveyed papers use either hamming distance and fractional hamming distance or jaccard index for measuring the properties of the PUFs. The usage of one over the other is debatable although Xiong et al claims jaccard index to be a better metric because it takes into consideration only the bits that flip as opposed to fractional hamming distance which weighs all bits of the CRP equally.

@@

\textit{Uniqueness.} Uniqueness in PUFs is measured by comparing responses for the same challenges on different PUF instances. Greater difference in responses indicates stronger uniqueness.  The comparison is measured using either hamming distance(needs to be high) or the jaccard index(needs to be low). All surveyed work proves the uniqueness through experimentation on PUF instances of different DRAM modules, Sutar et al. goes a step further to analyze the uniqueness of different PUF instances within the same DRAM module.

@@

\begin{figure}[!hbt]
    % Center the figure.
    \begin{center}
    % Include the eps file, scale it such that it's width equals the column width. You can also put width=8cm for example...
    \includegraphics[scale=0.5]{figs/enrollment.png}
    \setlength{\belowcaptionskip}{-20pt}
    % Create a subtitle for the figure.
    \caption{Enrollment}
    % Define the label of the figure. It's good to use 'fig:title', so you know that the label belongs to a figure.
    \label{fig:enroll}
    \end{center}
\end{figure}

@@

\textit{One Way Authentication.} The flowchart in Fig.\ref{fig:one_way} depicts the one-way authentication process. In short, the server matches the device's remeasured response against the CRP database. If matches, the device is authenticated

\begin{figure}[!hbt]
    % Center the figure.
    \begin{center}
    % Include the eps file, scale it such that it's width equals the column width. You can also put width=8cm for example...
    \includegraphics[scale=0.5]{figs/authentication.png}
    \setlength{\belowcaptionskip}{-20pt}
    % Create a subtitle for the figure.
    \caption{1-way Authentication}
    % Define the label of the figure. It's good to use 'fig:title', so you know that the label belongs to a figure.
    \label{fig:one_way}
    \end{center}
\end{figure}

\textit{Mutual Authentication.} This is an extension of the 1-way authentication which requires storage of CRPs in form of addresses of flipped bits rather than the bit pattern itself. The device mixes random addresses with the flipped addresses before sending the response to the server. This is done to test the knowledge of the server. The server verifies the device by counting the number of matched addresses and checking it against a threshold. The server then filters out the known flip addresses from the response and sends it to the device. The device verifies the server's response by matching it with its own response.

In both authentication protocols described above, it is important to note that:
\begin{itemize}
    \item A CRP is never used twice to prevent replay attacks. Hence a large CRP database or a strong PUF is required for continued authentication.
    \item For fast authentication it is critical to pick a CRP with the fastest possible parameters(low refresh pause interval). This requires knowledge of the PUF characteristics beforehand.
\end{itemize}

@@

\textit{Parameter Reconfigurability.} The main objective of parameter reconfigurability is to increase the size of the CRP database. For this, different parameters that can change the response behaviour of the PUF have to be recognized.

Sutar et al in their work show the reconfigurability of the refresh pause interval in order to obtain new challenge-response pairs on the same PUF instance. A similar experiment is also conducted by Xiong et al and Schaller et al. After the CRPs of a particular refresh pause interval have been exhausted, the interval can be increased to a higher one depending on the number of “new” bit flips. Both these works also propose increasing the size of the PUF instance, in order to maintain the low refresh pause interval at the cost of DRAM space exhaustion.

@@

With these parameters at disposal, one can reconfigure the PUF instance to generate new sets of CRP pairs and thereby increase the size of the CRP database.

\textit{Characterization:} For proper usage of the reconfiguration parameters, the challenge response behaviour must be analyzed before the authentication can take place. For achieving this the DRAM module has to be \textit{characterized}. In simple terms, characterization is the process of trying all kinds of combinations of parameters and observing the difference in responses.\cite{SUT18} After such an exhaustive experimentation the best suitable parameters are selected and the CRP database is generated accordingly.

@@

When a PUF cannot generate a large CRP database, it is used with a Fuzzy Extractor a.k.a Helper Data System \cite{HDS15} for secure key storage. A fuzzy extractor corrects errors in the PUF response \cite{SUT18} and enables regeneration of a cryptographic key. This way, the reproduced key can be used to setup a secure channel through symmetric protocols. Once a secure channel is established, authentication can be done through asymmetric protocols. A general construction of a fuzzy extractor contains:

\begin{itemize}
    \item \textit{Enrollment Algorithm:} The enrollment algorithm takes as input a PUF response \(X\) and a random cryptographic key \(k\) and outputs helper data \(w\).
    \item \textit{Reconstruction Algorithm:}  The reconstruction algorithm takes as input a fresh and noisy measurement of the PUF response \(X'\), and helper data \(w\), and outputs a reconstructed key \(k'\). If the noise in the PUF response is within a threshold \( k' = k\)
\end{itemize}

@@

One of the earliest implementations of fuzzy extractors is the Code Offset Method. \cite{COM} COM uses an error correcting code at the core of its construction. The helper data \(w = X \oplus Enc(k)\), where \(Enc()\) is the encoding function of the error correcting code. The reconstructed key \(k' = Dec(X' \oplus w)\), where \(Dec()\) is the decoding function of the error correcting code. The important point to note here is that \(w\) plays the role of a syndrome. The error correcting code extracts the key \(k\) by correcting errors in \(X'\). This is clear when the formula is rewritten as \(k' = Dec(X' \oplus X \oplus Enc(k))\). Any remaining bits of the \(X' \oplus X\), if there are any, is corrected by \(Dec()\)

Using such a fuzzy extractor would also require sufficient entropy in the input \(X\). Schaller et al overcomes this problem by selecting equal number of \textbf{F} cells and \textbf{S} cells. The entropy in the input is required so that no information is leaked through the helper data, \(w\).

Another version of the COM scheme, Syndrome-COM or SCOM,\cite{COM} leverages syndrome decoding to use the input, \(X\), itself as the cryptographic key. This is done by setting the helper data to the syndrome of \(X\). During reconstruction, The expression as shown in listings below evaluates to the syndrome of the error vector which is consequently corrected by the decoder.

\begin{algorithm}
\caption{Enroll}
\begin{algorithmic}[1]
    \STATE Measure \(X \in {0,1}^{n}\)
    \STATE Compute helper data \(W = SynX\).
    \STATE Publicly store \(W\)
\end{algorithmic}
\end{algorithm}

\begin{algorithm}
    \caption{Reconstruct}
    \begin{algorithmic}[1]
        \STATE Read \(W\)
        \STATE Measure \(X' \in {0,1}^{n}\)
        \STATE \(\hat{\text{X}} = X' \oplus SynDec(W \oplus Syn(X'))\)
    \end{algorithmic}
\end{algorithm}

In this scheme the helper data \(w = Syn(X)\), where \(Syn()\) is the syndrome function. The reconstructed key \(k' = SynDec(w \oplus Syn(X'))\).

@@

Other constructions, of HDSs make use of concatenation of two linear codes. \cite{HDS15}

\[C_{1} = [n_{1}, k_{1}, d_{1}] \quad\text{and}\quad C_{2} = [n_{2}, k_{2}, d_{2}]\]

In this construction, Code \(C_{1}\) needs to maintain a high entropy, which directly translates into high cardinality or high \(k_{1}\). And, Code \(C_{2}\) needs to have a high error correction rate which directly translates to higher distance, \(d_{2}\).

\section{Conclusion}
This paper highlights why and how PUFs fulfill the demands of hardware-intrinsic authentication. It explains the well known methods of implementing PUFs on DRAMs and the advantages of DRAM PUFs. It outlines the objectives an implementation of PUF has to meet in order to be fit for use in authentication. It summarizes a simple lightweight protocol that can meet strict resource constraints for embedded processors. It also dives into the details of using PUFs for secure key storage using helper data systems (HDS) and summarizes different constructions for HDS.

@@

@@
